---
title: "Difference in Difference"
subtitle: "2 x 2 Design"
output: 
  html_document:
    css: style_do.css
    number_sections: yes
    toc: yes
    toc_depth: '3'
    toc_float: yes
    theme: spacelab #"default", "bootstrap", "cerulean", "cosmo", "darkly", "flatly",
                 # "journal", "lumen", "paper", "readable", "sandstone", "simplex",
                 # "spacelab", "united", "yeti"
    df_print: paged
    code_folding: show
#date: '2022-03-30'
date: 'Last Update: `r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center")
library(foreach)
library(magrittr)
library(ggplot2)
library(kableExtra)
color_main <- scales::viridis_pal(option = "C")(1)
```



## Model

- $t = 1, 2, 3$.
- $r = 1 (= \infty), 2, 3$.
- Each group has 100 units.
- $Y_{it}(r) = Y_{it}(\infty) + \tau_t(r) + \epsilon_{it}(r), \epsilon_{it}(r) \sim N(0, 1)$.
- Common trend holds in the finite sample: $Y_{it}(\infty) = 0$.
- Common trend holds in the limit: $Y_{it}(\infty) = \epsilon_{it}(\infty), \epsilon_{it}(\infty) \sim N(0, 1)$.


## Heterogenous treatment effect

| $\tau_t(r)$ | $r = 1$ | $r = 2$ | $r = 3$ |
|-------------|---------|---------|---------|
| $t = 1$     | 1       | 2       | 3       |
| $t = 2$     | 4       | __5__   | 6       |
| $t = 3$     | 7       | __8__   | __9__   |


## Set parameters

- Common trend holds in the finite sample

```{r}
set.seed(1)
N_g <- 100
T <- 3
G <- T
N <- G * N_g
tau <- 
  tidyr::expand_grid(
    t = 1:T,
    g = 1:G
  ) %>%
  dplyr::mutate(
    tau = 1:length(t)
  )
```

## Make data


```{r}
df <-
  tidyr::expand_grid(
    t = 1:T,
    i = 1:N
  ) %>%
  dplyr::mutate(
    g = ceiling(i / 100)
  ) %>%
  dplyr::left_join(
    tau,
    by = c("g", "t")
  )
```

## Make data

```{r}
df <-
  df %>%
  dplyr::mutate(
    y_0 = 0,
    y_1 = tau + rnorm(length(t))
  ) %>%
  dplyr::mutate(
    d = (g > 1) * (t >= g),
    y = y_1 * d + y_0 * (1 - d)
  )
```


## Pooled OLS

```{r}
model_ols <- lm(
    data = df,
    formula = y ~ I((g == 2)&(t == 2)) + I((g == 2) & (t == 3)) + I((g == 3) & (t == 3)) + as.factor(g) + as.factor(t)
  )
modelsummary::modelsummary(model_ols, coef_omit = "as.factor")
```

## Two-way fixed effects

```{r}
model_twfe <- lfe::felm(
    data = df,
    formula = y ~ I((g == 2)&(t == 2)) + I((g == 2) & (t == 3)) + I((g == 3) & (t == 3)) | i + t 
  )
modelsummary::modelsummary(model_twfe)
```


## Separate DID for $\tau_2(2)$

```{r}
did_g2t2 <- 
  df %>%
  dplyr::filter(
    (g == 1 & t <= 2) |
      (g == 2 & t <= 2) 
  ) %>%
  lfe::felm(
    data = .,
    formula = y ~ d | g + t
  )
modelsummary::modelsummary(did_g2t2)
```

## Separate DID for $\tau_3(2)$

```{r}
did_g2t3 <- 
  df %>%
  dplyr::filter(
    (g == 1 & (t == 1 | t == 3)) |
      (g == 2 & (t == 1 | t == 3)) 
  ) %>%
  lfe::felm(
    data = .,
    formula = y ~ d | g + t
  )
modelsummary::modelsummary(did_g2t3)
```

## Separate DID for $\tau_3(3)$

```{r}
did_g3t3 <- 
  df %>%
  dplyr::filter(
    (g == 1) |
      (g == 3)
  ) %>%
  lfe::felm(
    data = .,
    formula = y ~ d | g + t
  )
modelsummary::modelsummary(did_g3t3)
```


## Set parameters

- Common trend holds only in the limit.

```{r}
set.seed(1)
N_g <- 100
T <- 3
G <- T
N <- G * N_g
tau <- 
  tidyr::expand_grid(
    t = 1:T,
    g = 1:G
  ) %>%
  dplyr::mutate(
    tau = 1:length(t)
  )
```

## Make data

```{r}
df <-
  tidyr::expand_grid(
    t = 1:T,
    i = 1:N
  ) %>%
  dplyr::mutate(
    g = ceiling(i / 100)
  ) %>%
  dplyr::left_join(
    tau,
    by = c("g", "t")
  ) 
```

## Make data

```{r}
df <-
  df %>%
  dplyr::mutate(
    y_0 = rnorm(length(t)),
    y_1 = tau + rnorm(length(t))
  ) %>%
  dplyr::mutate(
    d = (g > 1) * (t >= g),
    y = y_1 * d + y_0 * (1 - d)
  )
```


## Pooled OLS

```{r}
model_ols <- lm(
    data = df,
    formula = y ~ I((g == 2)&(t == 2)) + I((g == 2) & (t == 3)) + I((g == 3) & (t == 3)) + as.factor(g) + as.factor(t)
  )
modelsummary::modelsummary(model_ols, coef_omit = "as.factor")
```

## Two-way fixed effects

```{r}
model_twfe <- lfe::felm(
    data = df,
    formula = y ~ I((g == 2)&(t == 2)) + I((g == 2) & (t == 3)) + I((g == 3) & (t == 3)) | i + t 
  )
modelsummary::modelsummary(model_twfe)
```



## Separate DID for $\tau_2(2)$

```{r}
did_g2t2 <- 
  df %>%
  dplyr::filter(
    (g == 1 & t <= 2) |
      (g == 2 & t <= 2) 
  ) %>%
  lfe::felm(
    data = .,
    formula = y ~ d | g + t
  )
modelsummary::modelsummary(did_g2t2)
```

## Separate DID for $\tau_3(2)$

```{r}
did_g2t3 <- 
  df %>%
  dplyr::filter(
    (g == 1 & (t == 1 | t == 3)) |
      (g == 2 & (t == 1 | t == 3)) 
  ) %>%
  lfe::felm(
    data = .,
    formula = y ~ d | g + t
  )
modelsummary::modelsummary(did_g2t3)
```

## Separate DID for $\tau_3(3)$

```{r}
did_g3t3 <- 
  df %>%
  dplyr::filter(
    (g == 1 & (t == 1 | t == 3)) |
      (g == 3 & (t == 1 | t == 3))
  ) %>%
  lfe::felm(
    data = .,
    formula = y ~ d | g + t
  )
modelsummary::modelsummary(did_g3t3)
```

## Separate DID for $\tau_2(2)$ (using other treated groups as control)

```{r}
did_g2t2 <- 
  df %>%
  dplyr::filter(
    (g == 1 & t <= 2) |
      (g == 2 & t <= 2) |
      (g == 3 & t <= 2)
  ) %>%
  lfe::felm(
    data = .,
    formula = y ~ d | g + t
  )
modelsummary::modelsummary(did_g2t2)
```

## Separate DID for $\tau_3(2)$ (using other treated groups as control)

```{r}
did_g2t3 <- 
  df %>%
  dplyr::filter(
    (g == 1 & t <= 3) |
      (g == 2 & (t == 1 | t == 3)) |
      (g == 3 & (t <= 2))
  ) %>%
  lfe::felm(
    data = .,
    formula = y ~ d | g + t
  )
modelsummary::modelsummary(did_g2t3)
```

## Separate DID for $\tau_3(3)$ (using other treated groups as control)

```{r}
did_g3t3 <- 
  df %>%
  dplyr::filter(
    (g == 1 & t <= 3) |
      (g == 2 & t == 1) |
      (g == 3)
  ) %>%
  lfe::felm(
    data = .,
    formula = y ~ d | g + t
  )
modelsummary::modelsummary(did_g3t3)
```

## Testing pre-trend

```{r}
model_full <- lm(data = df, formula = y ~ -1 + as.factor(g):as.factor(t))
modelsummary::modelplot(model_full)
```

## Testing pre-trend

```{r}
anova(model_ols, model_full)
```

