---
title: "Cluster-Robust Standard Errors"
output: 
  html_document:
    css: style_do.css
    number_sections: yes
    toc: yes
    toc_depth: '3'
    toc_float: yes
    theme: "spacelab" #"default", "bootstrap", "cerulean", "cosmo", "darkly", "flatly",
                 # "journal", "lumen", "paper", "readable", "sandstone", "simplex",
                 # "spacelab", "united", "yeti"
    highlight: "pygments" #"default", "tango", "pygments", "kate", "monochrome",
                 # "espresso", "zenburn", "haddock", "textmate"
    df_print: paged
    code_folding: show
date: 'Last Update: `r Sys.Date()`'
---

We will introduce robust inference for cluster correlation using the `clubSandwich` package and the `fwildclusterboot` package.

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center")
library(ggplot2)
color_main <- scales::viridis_pal(option = "C")(1)
```

In addition to `set.seed()`, you need to fix `dqrng::dqset.seed()` in `fwildclusterboot`.

The function `return_dgp_cluster` defined in `function_inference` generates `N` observations for `N_C` clusters, and generates `X` following the normal distribution `N(0,1)`, `e` following the normal distribution `N(0,0.1)`, and `Y = 0.5 + X * tau_C + e` by adding cluster-specific normal errors to `e`. Here, `tau_C` gives the value divided from `[-2,2]` to each cluster.

Because the sample size is common for each cluster, the average treatment effect for the entire population is `0`.

# The Case of Quite a Few Clusters

First, set `N_C` to `100`. Using `clubSandwich::vcov` and `clubSandwich::coef_test` for the linear regression output of `lm`, we can obtain the estimated statistics based on cluster-robust variance estimation.

In this case, `vcov` can choose the degree of freedom correction for variance estimation. `CR0` is the one without correction, and `CR3` is the one by jackknife variance estimation.

```{r}
set.seed(1)
dqrng::dqset.seed(1)
N <- 50
N_C <- 100
N_M <- 5000
N_B <- 9999
```

```{r}
data <- 
  AppliedEconometrics::return_dgp_cluster(
    N = N,
    N_C = N_C
  )
```

```{r}
lm_fit <- 
  lm(
    Y ~ X,
    data = data
  )
```

```{r}
cat(
  "Heteroskedasticity-robust t-statistics", 
  summary(lm_fit)$coefficients[2,3], 
  "\n"
)

vc_cr0 <- 
  clubSandwich::vcovCR(
    lm_fit,
    cluster = data$C,
    type = "CR0"
  )

vc_cr3 <- 
  clubSandwich::vcovCR(
    lm_fit,
    cluster = data$C,
    type = "CR3"
  )

cat(
  "Cluster-robust t-statistics, CR0", 
  clubSandwich::coef_test(
    lm_fit,
    vc_cr0,
    coefs = "X"
  )$tstat, 
  "\n"
)

cat(
  "Cluster-robust t-statistics, CR3", 
  clubSandwich::coef_test(
    lm_fit,
    vc_cr3,
    coefs = "X"
  )$tstat, 
  "\n"
)
```

The function `fwildclusterboot::boottest` gives the estimated statistics based on wild cluster-bootstrap.

```{r}
p_val_w <- 
  fwildclusterboot::boottest(
    object = lm_fit,
    clustid = "C",
    param = "X",
    B = N_B,
    type = "webb"
  )$p_val
cat("Wild cluster-bootstrap p-value", p_val_w, "\n")
```

When `N_C` is `100`, the result of `5000` simulations is as follows.

```{r}
t_w <- 
  rep(
    NA,
    N_M
  )
t_cr0 <- 
  rep(
    NA,
    N_M
  )
t_cr3 <- 
  rep(
    NA,
    N_M
  )
p_cr_w <- 
  rep(
    NA,
    N_M
  )
for (
  i in 1:N_M
) {
  data <- 
    AppliedEconometrics::return_dgp_cluster(
      N = N,
      N_C = N_C
    )

  lm_fit <- 
    lm(
      Y ~ X,
      data = data
    )
  
  t_w[i] <- summary(lm_fit)$coefficients[2,3]

  vc_cr0 <- 
    clubSandwich::vcovCR(
      lm_fit,
      cluster = data$C,
      type = "CR0"
    )
  
  vc_cr3 <- 
    clubSandwich::vcovCR(
      lm_fit,
      cluster = data$C,
      type = "CR3"
    )

  t_cr0[i] <- 
    clubSandwich::coef_test(
      lm_fit,
      vc_cr0,
      coefs = "X"
    )$tstat
  
  t_cr3[i] <- 
    clubSandwich::coef_test(
      lm_fit,
      vc_cr3,
      coefs = "X"
    )$tstat

  invisible(
    capture.output(
      boot_lm <- 
        fwildclusterboot::boottest(
          object = lm_fit,
          clustid = "C",
          param = "X",
          B = N_B,
          type = "webb"
        )
      )
    )
    
  invisible(
    capture.output(
      p_cr_w[i] <- boot_lm$p_val
    )
  )
}
```

The (false) rejection probability based on heteroskedasticity-robust variance estimation:

```{r}
mean(abs(t_w) >= 1.965)
```

The (false) rejection probability based on cluster-robust variance estimation, CR0:

```{r}
mean(abs(t_cr0) >= 1.965)
```

The (false) rejection probability based on cluster-robust variance estimation, CR3:

```{r}
mean(abs(t_cr3) >= 1.965)
```

The (false) rejection probability based on wild-cluster bootstrap:

```{r}
mean(p_cr_w < 0.05)
```

# The Case of Quite a Few Clusters

When `N_C` is `10`, it is difficult to say that the number of clusters is sufficient for cluster-robust estimation.

In addition, the default `type=rademacher` of `boottest` gives a warning that there are limited values of the statistics when the number of clusters is around `10`. If you use `type=webb`, it gives weights of  `-(1/2)^{1/2}`,`-(3/2)^{1/2}`, `(1/2)^{1/2}`, `(3/2)^{1/2}`, in addition to `-1` and `1`.

[Webb, M. D. 2014. Reworking wild bootstrap based inference for clustered errors.
Queen’s University, Department of Economics, Working Paper No. 1315](https://ideas.repec.org/p/qed/wpaper/1315.html)

```{r}
set.seed(1)
dqrng::dqset.seed(1)
N <- 50
N_C <- 10
N_M <- 5000
N_B <- 9999
```

```{r}
data <- 
  AppliedEconometrics::return_dgp_cluster(
    N = N,
    N_C = N_C
  )
```

```{r}
lm_fit <- 
  lm(
    Y ~ X,
    data = data
  )

boot_lm <- 
  fwildclusterboot::boottest(
    object = lm_fit, 
    clustid = "C",
    param = "X",
    B = N_B
  )

summary(boot_lm)$statistic
```
。
When `N_C` is `10`, the result of `5000` simulations is as follows.

```{r}
t_w <- 
  rep(
    NA,
    N_M
  )
t_cr0 <- 
  rep(
    NA,
    N_M
  )
t_cr3 <- 
  rep(
    NA,
    N_M
  )
p_cr_w <- 
  rep(
    NA,
    N_M
  )

for (
  i in 1:N_M
  ) {

  data <- 
    AppliedEconometrics::return_dgp_cluster(
      N = N,
      N_C = N_C
    )

  lm_fit <- 
    lm(
      Y ~ X,
      data = data
    )

  t_w[i] <- summary(lm_fit)$coefficients[2,3]

  vc_cr0 <- 
    clubSandwich::vcovCR(
      lm_fit,
      cluster = data$C,
      type = "CR0"
    )

  vc_cr3 <- 
    clubSandwich::vcovCR(
      lm_fit,
      cluster = data$C,
      type = "CR3"
    )

  t_cr0[i] <- 
    clubSandwich::coef_test(
      lm_fit,
      vc_cr0,
      coefs = "X"
    )$tstat

  t_cr3[i] <- 
    clubSandwich::coef_test(
      lm_fit,
      vc_cr3,
      coefs = "X"
    )$tstat

  invisible(
    capture.output(
      boot_lm <- 
        fwildclusterboot::boottest(
          object = lm_fit,
          clustid = "C",
          param = "X",
          B = N_B,
          type = "webb"
        )
      )
    )
  invisible(
    capture.output(
      p_cr_w[i] <- boot_lm$p_val
    )
  )
}
```

The (false) rejection probability based on heteroskedasticity-robust variance estimation:

```{r}
mean(abs(t_w) >= 1.965)
```

The (false) rejection probability based on cluster-robust variance estimation, CR0:

```{r}
mean(abs(t_cr0) >= 1.965)
```

The (false) rejection probability based on cluster-robust variance estimation, CR3:

```{r}
mean(abs(t_cr3) >= 1.965)
```

The (false) rejection probability based on wild-cluster bootstrap:

```{r}
mean(p_cr_w < 0.05)
```

# The Case of Extremely Few Clusters

When `N_C` is `5`, the result of `5000` simulations is as follows.

```{r}
set.seed(1)
dqrng::dqset.seed(1)
N <- 100
N_C <- 5
N_M <- 5000
N_B <- 9999
```

```{r}
t_w <- 
  rep(
    NA,
    N_M
  )
t_cr0 <- 
  rep(
    NA,
    N_M
  )
t_cr3 <- 
  rep(
    NA,
    N_M
  )
p_cr_w <-
  rep(
    NA,
    N_M
  )

for (
  i in 1:N_M
  ) {
  
  data <- 
    AppliedEconometrics::return_dgp_cluster(
      N = N,
      N_C = N_C
    )

  lm_fit <- 
    lm(
      Y ~ X,
      data = data
    )

  t_w[i] <- summary(lm_fit)$coefficients[2,3]

  vc_cr0 <- 
    clubSandwich::vcovCR(
      lm_fit,
      cluster = data$C,
      type = "CR0"
    )

  vc_cr3 <- 
    clubSandwich::vcovCR(
      lm_fit,
      cluster = data$C,
      type = "CR3"
    )

  t_cr0[i] <- 
    clubSandwich::coef_test(
      lm_fit,
      vc_cr0,
      coefs = "X"
    )$tstat
  
  t_cr3[i] <- 
    clubSandwich::coef_test(
      lm_fit,
      vc_cr3,
      coefs = "X"
    )$tstat

  invisible(
    capture.output(
      boot_lm <- 
        fwildclusterboot::boottest(
          object = lm_fit,
          clustid = "C",
          param = "X",
          B = N_B,
          type = "webb"
        )
      )
    )
  invisible(
    capture.output(
      p_cr_w[i] <- boot_lm$p_val
    )
  )
}
```

The (false) rejection probability based on heteroskedasticity-robust variance estimation:

```{r}
mean(abs(t_w) >= 1.965)
```

The (false) rejection probability based on cluster-robust variance estimation, CR0:

```{r}
mean(abs(t_cr0) >= 1.965)
```

The (false) rejection probability based on cluster-robust variance estimation, CR3:

```{r}
mean(abs(t_cr3) >= 1.965)
```

The (false) rejection probability based on wild-cluster bootstrap:

```{r}
mean(p_cr_w < 0.05)
```
