---
title: "Non-compliance"
subtitle: "Two-sided"
output: 
  html_document:
    css: style_do.css
    number_sections: yes
    toc: yes
    toc_depth: '3'
    toc_float: yes
    theme: "spacelab" #"default", "bootstrap", "cerulean", "cosmo", "darkly", "flatly",
                 # "journal", "lumen", "paper", "readable", "sandstone", "simplex",
                 # "spacelab", "united", "yeti"
    highlight: "pygments" #"default", "tango", "pygments", "kate", "monochrome",
                 # "espresso", "zenburn", "haddock", "textmate"
    df_print: paged
    code_folding: show
date: 'Last updated: `r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center")
library(AppliedEconometricsCode)
library(foreach)
library(magrittr)
library(ggplot2)
library(kableExtra)
color_main <- scales::viridis_pal(option = "C")(1)
```

Here, we implement the estimation methods for the treatment assignment effect and the local average treatment effect in the case of "3 two-sided non-compliers" in the chapter on "Non-compliers". The functions called from `AppliedEconometrics` are defined in `R/functions_noncompliance.R`. 

First, we generate the potential outcomes of the sample, and then, we generate the treatment vector and the observed outcomes. At this time, we consider a situation where the average treatment effect differs for each compliance type. Based on the realized treatment assignment vector, treatment receipt vector, and observed outcomes, we estimate the treatment receipt and the treatment assignment effect on the outcome. Next, we use those estimates to estimate the local average treatment effect.

In the case of two-sided non-compliers, the absence of defiers is necessary to estimate the local average treatment effect. To verify this point, we generate data with and without the presence of defiers and compare the estimation results.

# Generating simulation data

## Setting constants and parameters

First, we set the seed, the number of individuals `N`, the number of draws in the Monte Carlo simulation `R`, and the size of the treatment group `N_1`. The set of compliance types is given by `G`.

```{r}
set.seed(1)
N <- 1000
N_1 <- 500
G <- 
  c(
    "nt",
    "co",
    "at",
    "df"
  )
```

Next, we set the average treatment effect on the outcome `tau_population` in the population.

```{r}
tau_population <- 
  c(
    1, 
    2, 
    3, 
    4
  )
```


## Generating data

The above constants and parameters are used to generate the potential outcomes of the sample. This operation is a single random draw. The finite-sample average treatment effect `tau` is also calculated for the random draw. The compliance types are represented by `g`.

```{r}
outcome_potential <-
  1:length(G) %>%
  purrr::map(
    function (n) {
      df <-
        tibble::tibble(
          g = G[n],
          y_0 = 
            rnorm(
              N / length(G), 
              mean = 0, 
              sd = 1
            ),
          y_1 = 
            rnorm(
              N / length(G), 
              mean = tau_population[n], 
              sd = 1
            )
        )
      return(df)
    }
  ) %>%
  dplyr::bind_rows()

tau <-
  outcome_potential %>%
  dplyr::group_by(g) %>%
  dplyr::summarise(
    tau = 
      mean(
        y_1 - y_0
      )
  ) %>%
  dplyr::ungroup() 

outcome_potential %>%
  head() %>%
  kbl() %>%
  kable_styling()

outcome_potential %>%
  modelsummary::datasummary_skim()

tau %>%
  kbl() %>%
  kable_styling()
```


Now, in this setting, there are defiers. As we will see in the following analysis, the average treatment effect cannot be accurately estimated with this data. Therefore, let's also generate data without any defiers.

```{r}
outcome_potential_nodefier <-
  outcome_potential %>%
  dplyr::filter(g != "df")
```


Next, we generate the observed data. To do this, we first generate the treatment assignment vector `z`. Based on this treatment assignment vector and the compliance type, we generate the treatment receipt vector `d`. This operation corresponds to a single randomized experiment. Note that the realized data does not include the compliance type `g`. We prepare data both with and without defiers.

```{r}
data_realized <-
  generate_data_noncompliance_twoside(
    outcome_potential = outcome_potential,
    N_1 = N_1,
    seed = 1
  ) 
data_realized %>%
  head() %>%
  kbl() %>%
  kable_styling()
data_realized %>%
  modelsummary::datasummary_skim()
```

```{r}
data_realized_nodefier <-
  generate_data_noncompliance_twoside(
    outcome_potential = outcome_potential_nodefier,
    N_1 = N_1,
    seed = 1
  ) 
data_realized_nodefier %>%
  head() %>%
  kbl() %>%
  kable_styling()
data_realized_nodefier %>%
  modelsummary::datasummary_skim()
```


# Analysis


## Manually calculating $\tilde{\tau}_D$

Based on the observed data, calculate the difference in the probability of receiving the treatment when it is assigned versus when it is not assigned. This serves as the estimated effect of treatment assignment on treatment receipt, denoted as $\tilde{\tau}_D$.

```{r}
tau_tilde_d <-
  data_realized %>%
  dplyr::summarise(
       d = 
        sum(d * z) / sum(z) - 
        sum(d * (1 - z)) / sum(1 - z)
    ) %>%
  dplyr::pull(d)
tau_tilde_d
```

Thus, in cases where defiers are present, the value of $\tilde{\tau}_D$ significantly differs from the share of compliers (`r 1/4`).

```{r}
tau_tilde_d_nodefier <-
  data_realized_nodefier %>%
  dplyr::summarise(
    d = 
      sum(d * z) / sum(z) - 
      sum(d * (1 - z)) / sum(1 - z)
  ) %>%
  dplyr::pull(d)
tau_tilde_d_nodefier
```

On the other hand, in cases without defiers, the value of $\tilde{\tau}_D$ is close to the share of compliers (`r 2/3`).

## Estimating $\tilde{\tau}_D$ using regression analysis

The same estimate can be obtained by regression analysis.

```{r}
data_realized %>%
  lm(
    data = .,
    formula = d ~ z
  ) %>%
  modelsummary::modelsummary(
    fmt = 6
  )
```


```{r}
data_realized_nodefier %>%
  lm(
    data = .,
    formula = d ~ z
  ) %>%
  modelsummary::modelsummary(
    fmt = 6
  )
```

## Manually calculating $\tilde{\tau}_Y$

Based on the observed data, calculate the difference in the average of the outcomes. This serves as the estimated effect of treatment assignment on the outcome, denoted as $\tilde{\tau}_Y$.

```{r}
tau_tilde_y <-
  data_realized %>%
  dplyr::summarise(
    y = 
    sum(y * z) / sum(z) - 
    sum(y * (1 - z)) / sum(1 - z)
  ) %>%
  dplyr::pull(y)
tau_tilde_y
```

```{r}
tau_tilde_y_nodefier <-
  data_realized_nodefier %>%
  dplyr::summarise(
    y = 
      sum(y * z) / sum(z) - 
      sum(y * (1 - z)) / sum(1 - z)
  ) %>%
  dplyr::pull(y)
tau_tilde_y_nodefier
```

## Estimating $\tilde{\tau}_Y$ using regression analysis

The same estimate can be obtained by regression analysis.

```{r}
data_realized %>%
  lm(
    data = .,
    formula = y ~ z
  ) %>%
  modelsummary::modelsummary(
    fmt = 6
  )
```

```{r}
data_realized_nodefier %>%
  lm(
    data = .,
    formula = y ~ z
  ) %>%
  modelsummary::modelsummary(
    fmt = 6
  )
```

## Manually calculating $\tilde{\tau}_{Y, co} = \tau_{late}$

Next, calculate the estimated value of the local average treatment effect by dividing the treatment assignment effect on the outcome by the treatment assignment effect on treatment receipt.

```{r}
tau_late <-
  tau_tilde_y / 
  tau_tilde_d
tau_late
```

Thus, in cases where defiers are present, the value of $\tilde{\tau}_{Y, co}$ significantly differs from the average treatment effect on the outcome (`r 2`).

```{r}
tau_late_nodefier <-
  tau_tilde_y_nodefier / 
  tau_tilde_d_nodefier
tau_late_nodefier
```

On the other hand, in cases without defiers, the value of $\tilde{\tau}_{Y, co}$ is close to the average treatment effect on the outcome (`r 2`).


## Estimating $\tilde{\tau}_{Y, co} = \tau_{late}$ using two-stage least squares

The same estimate can be obtained by using two-stage least squares. Estimation of the estimation error is more efficient when regression analysis is used.

```{r}
data_realized %>%
  estimatr::iv_robust(
    data = .,
    formula = y ~ d | z
  ) %>%
  modelsummary::modelsummary(
    fmt = 6
  )
```


```{r}
data_realized_nodefier %>%
  estimatr::iv_robust(
    data = .,
    formula = y ~ d | z
  ) %>%
  modelsummary::modelsummary(
    fmt = 6
  )
```


























