---
title: "Randomized Experiment"
subtitle: "Clustered Randomized Experiments"
output: 
  html_document:
    css: style_do.css
    number_sections: yes
    toc: yes
    toc_depth: '3'
    toc_float: yes
    theme: spacelab #"default", "bootstrap", "cerulean", "cosmo", "darkly", "flatly",
                 # "journal", "lumen", "paper", "readable", "sandstone", "simplex",
                 # "spacelab", "united", "yeti"
    df_print: paged
    code_folding: show
#date: '2022-03-30'
date: 'Last Update: `r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center")
library(foreach)
library(magrittr)
library(ggplot2)
library(kableExtra)
library(modelsummary)
color_main <- scales::viridis_pal(option = "C")(1)
```


Translate: Here, we implement the clustered randomized experiment. The functions called from `AppliedEconometricsCode` are defined in `R/functions_randomized.R`. First, we generate the potential outcomes of the sample, and then we generate the treatment vector and the realized outcomes. We consider two methods for estimating the average treatment effect: one that treats each cluster as an individual and calculates the average treatment effect ($\hat{\tau}_c$), and one that calculates the average treatment effect for the entire sample ($\hat{\tau}$). Finally, we calculate the cluster-robust standard error when estimating the average treatment effect using the data from the clustered randomized experiment.

# Generating simulated data

## Calculating the constants and parameters

We consider a clustered randomized experiment with two clusters. First, we fix the seed and set the number of clusters `G` and draw the number of individuals in each cluster `N_g` from a Poisson distribution. We set the total number of individuals `N`. We specify the number of clusters to which the treatment is assigned as `G_1`.


```{r}
set.seed(1)
G <- 100
N_g <- 
  rpois(
    G, 
    99
  ) + 
  1
N <- sum(N_g)
G_1 <- 50
```

Next, we draw the average treatment effect `tau_g_population` in each cluster from the absolute value of a standard normal distribution.

```{r}
tau_g_population <- 
  abs(
    rnorm(G)
  )
tau_g_population
```

The average treatment effect `tau_population` in the population is defined as the weighted average of `tau_g_population` using the number of individuals in each cluster.

```{r}
tau_population <- 
  sum(
    tau_g_population * N_g / 
    N
  )
tau_population
```

The population average treatment effect `tau_c_population` when each cluster is considered as an individual is defined as the simple average of `tau_g_population`.

```{r}
tau_c_population <- mean(tau_g_population)
tau_c_population
```

## Generating data

Based on the above constants and parameters, we generate the potential outcomes of the sample. This operation corresponds to one random draw.

```{r}
outcome_potential <-
  1:G %>%
  purrr::map(
    .,
    ~ tibble::tibble(
      g = .,
      y_0 = 
        rnorm(
          N_g[.], 
          mean = 0, 
          sd = 1
        ),
      y_1 = 
        rnorm(
          N_g[.], 
          mean = tau_g_population[.], 
          sd = 1
        )
    )
  ) %>%
  dplyr::bind_rows()

outcome_potential %>%
  head() %>%
  kbl() %>%
  kable_styling()

outcome_potential %>%
  modelsummary::datasummary_skim()
```

We also calculate the average treatment effect `tau_g` in each cluster in the sample.

```{r}
tau_g <-
  outcome_potential %>%
  dplyr::group_by(g) %>%
  dplyr::summarise(
    tau = 
      mean(
        y_1 - y_0
      )
  ) %>%
  dplyr::pull(tau)

tau_g
```

The average treatment effect `tau` in the sample is also calculated.

```{r}
tau <-
  outcome_potential %>%
  dplyr::summarise(
    tau = 
      mean(
        y_1 - y_0
      )
  ) %>%
  dplyr::pull(tau)

tau
```

We also calculate the average treatment effect `tau_c` in the sample.

```{r}
tau_c <- mean(tau_g)
tau_c
```

Next, we generate the observed data. First, we generate the treatment vector `z`. This operation corresponds to one random experiment. We then calculate the realized outcomes based on the realized treatment vector and potential outcomes. The resulting `data_realized` is the observed data.

```{r}
data_realized <-
  AppliedEconometricsCode::generate_data_cluster(
    outcome_potential = outcome_potential,
    G_1 = G_1,
    seed = 1
  )
data_realized %>%
  head() %>%
  kbl() %>%
  kable_styling()
```

# Analysis

## Estimating $\hat{\tau}_c$ using the difference in means

```{r}
tau_c_hat <-
  data_realized %>%
  dplyr::group_by(g) %>%
  dplyr::summarise(
    z = mean(z),
    y = mean(y)
    ) %>%
  dplyr::ungroup() %>%
  dplyr::summarise(
    tau = 
      sum(y * z) / 
      sum(z) - 
      sum(y * (1 - z)) / 
      sum(1 - z)
  )
tau_c_hat
```

## Calculating the standard error of $\hat{\tau}_c$ manually

```{r}
tau_c_hat_se <-
  data_realized %>%
  dplyr::group_by(g) %>%
  dplyr::summarise(
    z = mean(z),
    y = mean(y)
    ) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(z) %>%
  dplyr::summarise(
    G_z = length(y),
    variance = 
      sum(
        (y - mean(y))^2 / 
        (G_z - 1)
      )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::summarise(
    se = 
      sum(
        variance / 
        G_z
      ) %>% 
      sqrt()
  )
  
tau_c_hat_se
```

## Estiamting $\hat{\tau}_c$ using regression analysis

```{r}
tau_c_hat_regression <-
  data_realized %>%
  dplyr::group_by(g) %>%
  dplyr::summarise(
    z = mean(z),
    y = mean(y)
    ) %>%
  dplyr::ungroup() %>%
  lm(
    data = .,
    formula = y ~ z
  ) 
```

## The estimate of $\hat{\tau}_c$ 

```{r}
tau_c_hat_regression %>%
  modelsummary(
    fmt = 6
  )
```

## Estimating $\hat{\tau}$ using the difference in means

```{r}
tau_hat <-
  data_realized %>%
  dplyr::summarise(
    tau = 
      sum(y * z) / 
      sum(z) - 
      sum(y * (1 - z)) / 
      sum(1 - z)
  )
tau_hat
```

## Estimating $\hat{\tau}$ using regression analysis

```{r}
tau_hat_regression <-
  data_realized %>%
  dplyr::mutate(
    z = 
      z %>%
      as.integer()
    ) %>%
  lm(
    data = .,
    formula = y ~ z
  )
```

## The estimate of $\hat{\tau}$ 

```{r, warning = FALSE}
tau_hat_regression %>%
  modelsummary(
    vcov = ~ g, 
    fmt = 6
  )
```

## Calculating the standard error of $\hat{\tau}$ manually

```{r}
se_cluster_robust <-
  data_realized %>%
  dplyr::mutate(
    constant = 1,
    epsilon = tau_hat_regression$residuals
  ) 
  
term_1 <-
  se_cluster_robust %>%
  dplyr::select(
    constant, 
    z
  ) %>%
  as.matrix()
term_1 <- 
  crossprod(
    term_1, 
    term_1
  )
```

```{r}
term_2 <-
  se_cluster_robust %>%
  dplyr::group_split(g) %>%
  purrr::map(
    .,
    function(df) {
      df <-
        df %>%
        dplyr::mutate(
          z_epsilon = z * epsilon
        ) %>%
        dplyr::select(
          epsilon, 
          z_epsilon
        ) %>%
        dplyr::summarise_all(sum) %>%
        as.matrix()
      df <- 
        crossprod(
          df, 
          df
        )
    }
  ) %>%
  purrr::reduce(`+`)
```


```{r}
se_cluster_robust <-
  solve(
    term_1, 
    term_2
  ) %*% 
  solve(term_1)
se_cluster_robust %>%
  diag() %>%
  sqrt() 
```
