---
title: "Randomized Experiment"
subtitle: "Estimation of Average Treatment Effect by Difference in Means"
output: 
  html_document:
    css: style_do.css
    number_sections: yes
    toc: yes
    toc_depth: '3'
    toc_float: yes
    theme: "spacelab" #"default", "bootstrap", "cerulean", "cosmo", "darkly", "flatly",
                 # "journal", "lumen", "paper", "readable", "sandstone", "simplex",
                 # "spacelab", "united", "yeti"
    highlight: "pygments" #"default", "tango", "pygments", "kate", "monochrome",
                 # "espresso", "zenburn", "haddock", "textmate"
    df_print: paged
    code_folding: show
#date: '2022-03-30'
date: 'Last Update: `r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center")
library(foreach)
library(magrittr)
library(ggplot2)
library(kableExtra)
color_main <- scales::viridis_pal(option = "C")(1)
```

Here we implement the method of "6 Estimation of Average Treatment Effect". The functions called from `AppliedEconometricsCode` are defined in `R/functions_randomized.R`. First, we generate the potential outcomes of the sample, and then we generate the treatment vector and the realized outcomes. Based on the realized treatment vector and outcomes, we obtain the estimator of the average treatment effect based on the difference in means. Finally, we calculate the standard error of the estimator of the average treatment effect using Neyman's method.

# Generate simulation data

## Setting of constants and parameters

First, we set the seed and the number of units `N`, the number of draws in the Monte Carlo simulation `R`, and the size of the treatment group `N_1`.

```{r}
set.seed(1)
N <- 1000
R <- 1000
N_1 <- 500
```

Next, we set the average treatment effect `tau_population` in the population.

```{r}
tau_population <- 0.2
```

## Generate data

We generate the potential outcomes of the sample based on the above constants and parameters. This operation corresponds to one draw of random sampling. We also calculate the average treatment effect `tau` in the sample obtained by random draw.

```{r}
outcome_potential <-
  tibble::tibble(
    y_0 = 
      rnorm(
        N, 
        mean = 0, 
        sd = 1
      ),
    y_1 = 
      rnorm(
        N, 
        mean = tau_population, 
        sd = 1
      )
  )
tau <-
  outcome_potential %>%
  dplyr::summarise(
    tau = mean(y_1 - y_0)
  ) %>%
  dplyr::pull(tau)

outcome_potential %>%
  head() %>%
  kbl() %>%
  kable_styling()

outcome_potential %>%
  modelsummary::datasummary_skim()

tau
```

Next, we generate the observed data. To do this, we first generate the treatment vector `z`. This operation corresponds to one randomized experiment. Furthermore, we calculate the realized outcomes from the realized treatment vector and the potential outcomes. This part of the process is the same as in `randomization_fisher_pvalue.Rmd`.

```{r}
data_realized <-
  AppliedEconometricsCode::generate_data_randomized(
    outcome_potential = outcome_potential,
    N_1 = N_1,
    seed = 1
  )
data_realized %>%
  head() %>%
  kbl() %>%
  kable_styling()
data_realized %>%
  modelsummary::datasummary_skim()
```


# Analysis

## Calculation of the difference in means

We calculate the difference in means of the outcomes based on the observed data. This is the estimator of the average treatment effect.

```{r}
mean_1 <- 
  data_realized %>% 
  dplyr::filter(
    z == 1
  ) %>%
  dplyr::summarise(
    y = mean(y)
  ) %>%
  dplyr::pull(y)
mean_0 <- 
  data_realized %>% 
  dplyr::filter(
    z == 0
  ) %>%
  dplyr::summarise(
    y = mean(y)
  ) %>%
  dplyr::pull(y)
n_1 <-   
  data_realized %>% 
  dplyr::filter(
    z == 1
  ) %>%
  nrow()
n_0 <-   
  data_realized %>% 
  dplyr::filter(
    z == 0
  ) %>%
  nrow()
tau_hat <- mean_1 - mean_0
tau_hat
```

## Calculating the standard error

Next, we estimate the standard error of the average treatment effect using Neyman's method.

```{r}
var_1 <- 
  data_realized %>% 
  dplyr::filter(
    z == 1
  ) %>%
  dplyr::summarise(
    y = 
      sum(
        (y - mean_1)^2 / 
        (n_1 - 1)
      )
  ) %>%
  dplyr::pull(y)
var_0 <- 
  data_realized %>% 
  dplyr::filter(
    z == 0
  ) %>%
  dplyr::summarise(
    y = 
    sum(
      (y - mean_0)^2 / 
      (n_0 - 1)
    )
  ) %>%
  dplyr::pull(y)
var_hat <-
  var_1 / n_1 + 
  var_0 / n_0
var_hat
se_hat <- sqrt(var_hat)
se_hat
```

We summarize the estimation results in a table.

```{r}
data.frame(
  tau_hat,
  se_hat,
  tau,
  tau_population
) %>%
magrittr::set_rownames("Average treatment effect") %>%
kbl(
  col.names = c(
    "Estimate",
    "S.E.",
    "True (Sample)",
    "True (Population)"
  )
) %>%
kable_styling()
```

