---
title: "Regression Discontinuity"
subtitle: "Fuzzy RD Design"
output: 
  html_document:
    css: style_do.css
    number_sections: yes
    toc: yes
    toc_depth: '3'
    toc_float: yes
    theme: "spacelab" #"default", "bootstrap", "cerulean", "cosmo", "darkly", "flatly",
                 # "journal", "lumen", "paper", "readable", "sandstone", "simplex",
                 # "spacelab", "united", "yeti"
    highlight: "pygments" #"default", "tango", "pygments", "kate", "monochrome",
                 # "espresso", "zenburn", "haddock", "textmate"
    df_print: paged
    code_folding: show
date: 'Last update: `r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center")
library(AppliedEconometrics)
library(foreach)
library(magrittr)
library(ggplot2)
library(kableExtra)
library(modelsummary)
library(rdrobust)
color_main <- scales::viridis_pal(option = "C")(1)
```

Here, we implement the estimation method for fuzzy regression discontinuity design using the `rdrobust` package. The functions called from `AppliedEconometrics` are defined in `R/functions_rdrobust.R`. Below, we generate samples based on Lee (2008), set options for the `rdrobust` package, and explain the interpretation of the results.

# Generating Simulation Data

## Setting Constants and Parameters

First, fix the seed and set the number of individuals `N`.

```{r}
set.seed(1)
N <- 1000
```

## Generating Data

Next, we generate samples for the fuzzy RD design. The conditional expectation function used for data generation is the same as the one used in the `rdrobust` simulations, similar to Lee (2008). There are two differences from the sharp design:

- The potential outcome function is defined based on the treatment variable `d` with noncompliance, not the instrumental variable `z`.
- The effect size is increased to detect effects even with noncompliance.

The increased effect size of the induced model, i.e., ITT, is `0.24`, and since the compliance rate is `0.6`, the LATE is `0.4`. Here, the treatment variable is assumed to follow a probit model.

```{=latex}
\begin{align*}
 d =& 1\{\Psi^{-1}(0.3(1 + 0.5 s) + 0.4z) \leq \epsilon_i\}, \epsilon_i \sim N(0,1)\\
 z =& 1\{s \geq 0\}, \Psi(x) = P(\epsilon_i \leq x),\\
 y =& y^*(1)d + y^*(0)(1 - d)
\end{align*}
```

Generate Lee data and name it `dgp_lee`.

```{r}
dgp_lee <- 
  generate_dgp_lee_fuzzy(
    N = N
  )
```

# Analysis

## Estimation with the `rdrobust` package

Similar to the sharp design, by specifying the outcome variable `y` and the score variable `x` to the `rdrobust` function, a list of results for the ITT effect can be obtained.

```{r}
result <- 
  rdrobust::rdrobust(
    y = dgp_lee$y,
    x = dgp_lee$s,
    all = TRUE
  )
summary(result)
```

Similarly, by using the treatment variable `d` with noncompliance instead of the outcome variable, the effect of the instrumental variable on the treatment can be obtained.

```{r}
result <- 
  rdrobust::rdrobust(
    y = dgp_lee$d,
    x = dgp_lee$s,
    all = TRUE
  )
summary(result)

```

When specifying the treatment variable `d` with noncompliance in the `fuzzy` option, the output for the fuzzy RD design can be obtained.

```{r}
result <- 
  rdrobust::rdrobust(
    y = dgp_lee$y,
    x = dgp_lee$s, 
    fuzzy = dgp_lee$d,
    all = TRUE
  )
summary(result)
```

The ITT was `0.24`, and the compliance rate was `0.6`, resulting in a LATE of `0.4`. Consequently, the results obtained were consistent with the data generation process.


