
---
title: "Regression Discontinuity"
subtitle: "Regression Kink Design"
output: 
  html_document:
    css: style_do.css
    number_sections: yes
    toc: yes
    toc_depth: '3'
    toc_float: yes
    theme: "spacelab" #"default", "bootstrap", "cerulean", "cosmo", "darkly", "flatly",
                 # "journal", "lumen", "paper", "readable", "sandstone", "simplex",
                 # "spacelab", "united", "yeti"
    highlight: "pygments" #"default", "tango", "pygments", "kate", "monochrome",
                 # "espresso", "zenburn", "haddock", "textmate"
    df_print: paged
    code_folding: show
date: 'Last update: `r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center")
library(AppliedEconometrics)
library(foreach)
library(magrittr)
library(ggplot2)
library(kableExtra)
library(modelsummary)
color_main <- scales::viridis_pal(option = "C")(1)
```

## Set simulation parameters

```{r}
set.seed(1)
N <- 1000
```

## Simulate data

```{r}
df <-
  tibble::tibble(
    u = rnorm(N),
    v = rnorm(N) + 0.1 * u
  ) %>%
  dplyr::mutate(
    b = compute_b(v),
    y = compute_y(b, v, u)
  )
```

## Kink between v and b

```{r}
df %>% 
  ggplot(
    aes(
      x = v, 
      y = b
    )
  ) + 
  geom_point() + 
  theme_classic()
```

## Kink between v and y

```{r}
df %>% 
  ggplot(
    aes(
      x = v, 
      y = y
    )
  ) + 
  geom_point() + 
  theme_classic()
```

## Kink between v and y with a binned scatter plot

```{r}
rdrobust::rdplot(
  y = df$y, 
  x = df$v, 
  c = 0, 
  binselect = "espr", 
  x.label = "v", 
  y.label = "y"
)
```

## `rdrobust::rdrobust` does not accept a sharp RKD with a continuous treatment
- The following is the inference for $\lim_{v \downarrow 0} \frac{d \mathbb{E}[Y|V = v]}{\partial v} - \lim_{v \uparrow 0} \frac{d \mathbb{E}[Y|V = v]}{\partial v}$.
- Dividing by $\lim_{v \downarrow 0} b'(v) - \lim_{v \uparrow 0} b'(v) = 1 - 2 = -1$ gives roughly an estimate of 2.

```{r}
result <- 
  rdrobust::rdrobust(
    y = df$y, 
    x = df$v, 
    c = 0, 
    deriv = 1, 
    kernel = "triangular", 
    bwselect = "mserd", 
    all = "true"
  )
cbind(
  result$coef, 
  result$se
) %>% 
kbl() %>% 
kable_styling()
```
