---
title: "Observational Study"
subtitle: "Regular Assignment Mechanism"
output: 
  html_document:
    css: style_do.css
    number_sections: yes
    toc: yes
    toc_depth: '3'
    toc_float: yes
    theme: spacelab #"default", "bootstrap", "cerulean", "cosmo", "darkly", "flatly",
                 # "journal", "lumen", "paper", "readable", "sandstone", "simplex",
                 # "spacelab", "united", "yeti"
    df_print: paged
    code_folding: show
#date: '2022-03-30'
date: 'Last Update: `r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center")
library(foreach)
library(magrittr)
library(ggplot2)
library(kableExtra)
library(modelsummary)
color_main <- scales::viridis_pal(option = "C")(1)
```

Here, we implement methods in the observational study with regular assignment mechanism. We consider the following methods: imputing by a linear regression, using a weighing estimator, using a blocking estimator, and using a matching estimator.

# Generating simulation data

## Setting constants and parameters

First, we set the seed and the number of observations `N`, blocks for the blocking estimator `J`, and neighbors for the matching estimator `M`. We also set the true coefficients.

```{r}
set.seed(1)
N <- 1000
J <- 4
M <- 1
beta_t <- c(1, 0.1)
beta_c <- c(0, 0.05)
```

## Generating data

We generate the data for the observational study. We generate the covariate `x` from a standard normal distribution and the potential outcomes `y_0` and `y_1` from normal distributions with the true coefficients. 

```{r}
outcome <-
  tibble::tibble(
    x = rnorm(N),
    y_0 = beta_c[1] + beta_c[2] * x + rnorm(N),
    y_1 = beta_t[1] + beta_t[2] * x + rnorm(N)
  )
head(outcome)
```

We calculate the average treatment effect in the sample.

```{r}
tau_fs <- 
  outcome %>%
  dplyr::summarise(tau = mean(y_1 - y_0)) %>%
  dplyr::pull(tau)
tau_fs
```

We calculate the propensity score and assign the treatment based on the propensity score. Thus, the assignment mechanisms depends on the covariates.

```{r}
outcome <-
  outcome %>%
  dplyr::mutate(
    e = exp(x) / (1 + exp(x)),
    z = (runif(N) < e)
    )
head(outcome)
```

## Check the balance in the covariate

We check the balance in the covariate `x` between the treatment and control groups.

```{r, echo = FALSE, message = FALSE}
outcome %>%
  ggplot(
    aes(
      x = x,
      fill = z
    )
  ) +
  geom_histogram() +
  scale_fill_viridis_d() +
  theme_classic()
```


```{r}
outcome_observed <-
  outcome %>%
  dplyr::mutate(y = y_0 * (1 - z) + y_1 * z) %>%
  dplyr::select(y, z, x)
head(outcome_observed)
```

# Analysis

## Imputing by a linear regression

We estimate the average treatment effect by imputing the potential outcomes by a linear regression.

```{r}
result <-
  lm(
    data = outcome_observed,
    formula = y ~ z + z:x 
  )
result %>%
  modelsummary(fmt = 6)
```

```{r}
result <-
  c(
    y ~ z + z:exp(x),
    y ~ z + z:x^2,
    y ~ z + z:(x + x^2)
  ) %>%
  purrr::map(
    .,
    ~ lm(
      data = outcome_observed,
      formula = .
      )
  )
```


```{r}
result %>%
  modelsummary(fmt = 6)
```


## Using a weighing estimator

We first estimate the propensity score by a correctly specified logistic function.

```{r}
propensity_score <-
  glm(
    data = outcome_observed,
    formula = z ~ x,
    family = binomial("logit")
  )
propensity_score %>%
  modelsummary(fmt = 6)
```

```{r, eval = FALSE}
outcome_observed <-
  outcome_observed %>%
  dplyr::mutate(e = predict(propensity_score, type = "response"))
outcome_observed %>%
  ggplot(
    aes(
      x = x,
      y = e
    )
  ) +
  geom_line() +
  theme_classic()
```

```{r, echo = FALSE}
outcome_observed <-
  outcome_observed %>%
  dplyr::mutate(e = predict(propensity_score, type = "response"))
outcome_observed %>%
  ggplot(
    aes(
      x = x,
      y = e
    )
  ) +
  geom_line() +
  theme_classic()
```

```{r}
outcome_observed %>%
  dplyr::mutate(weight = (z / e + (1 - z) / (1 - e)) / length(z)) %>%
  dplyr::summarise(y = sum(z * y * weight) - sum((1 - z) * y * weight)) 
```

## Using a blocking estimator

```{r}
blocking_estimator <-
  outcome_observed %>%
  dplyr::mutate(j = dplyr::ntile(outcome_observed$e, J)) %>%
  dplyr::group_by(j) %>%
  dplyr::summarise(
    N_j = length(y),
    y = sum(y * z) / sum(z) - sum(y * (1 - z)) / sum(1 - z)
    ) %>%
  dplyr::ungroup()
blocking_estimator
```


```{r}
blocking_estimator %>%
  dplyr::summarise(y = sum(y * N_j) / sum(N_j))
```

## Using a matching estimator

```{r}
matched <-
  MatchIt::matchit(
    data = outcome_observed,
    formula = z ~ x,
    method = "nearest",
    distance = "glm",
    ratio = M
  )
 matched 
```

```{r}
matched_observed <-
  MatchIt::match.data(matched) 
matched_observed
```

```{r}
matched_observed %>%
  dplyr::group_by(subclass) %>%
  dplyr::summarise(y = sum(y * z) - sum(y * (1 - z))) %>%
  dplyr::ungroup() %>%
  dplyr::summarise(y = mean(y))
```
